---
title: "Visualize!"
author: "Mary Krouse"
date: "2026-07-27"
output: html_document
---

# Setup

```{r}
library(tidyverse)
library(ggplot2)
library(showtext)
library(dplyr)
library(pspearman)
```

```{r}
survey <- read_csv("answers_clean.csv") %>%
  mutate(
    Politics = fct_relevel(
      Politics, 
      "Far left", "Moderate left", "Center",
      "Moderate right", "Far right"
    ),
    Location = case_match(
      Location, 
      "Online or sent to me, and I'm taking this survey from within the United States" ~ "Online within US",
    "Online or sent to me, and I'm taking this survey from outside the United States" ~ "Online outside US"
    .default = Location
  )
)
```

## Contents

Intersecting Favorability Distributions (2)
- proportions of furry favorability for each bucket of trans/gay favorability
- color gradient to represent proportion
- hue to represent topic

Prediction Heatmap (1)
- proportions of predicted furry favorability for each bucket of true furry favorability
- color gradient to represent correctness vs incorrectness
- highlighting to represent correctness

Favorability Score Distributions (3)
- distributions of each favorability score topic
- height to represent count
- hue to represent topic

Question Distributions (66 + 9)
- distributions of each question
- height to represent count

Raw Favorability Distributions (3)
- distributions of raw favorability scores before categorical aggregation for each topic
- height to represent density
- hue to represent topic

## Universal Variables

```{r}
# image resolution

my_dpi = 600

# fonts

font_add_google("Roboto Slab", "roboto_slab")
showtext_auto()
showtext_opts(dpi = my_dpi)

# colors

yellow = "#FFC532"
green = "#65CB4B"
cyan = "#00C6AC"

pink = "#FF7E75"
pink_mild = "#F8A19B"
pink_soft = "#F0C3C0"
pink_faint = "#FAEBEA"
pink_dark = "#B30300"

lo_prop = "#FAFAFA"

pink_hi_prop = "#FC6055"
pink_hi_prop_txt = pink_dark
pink_lo_prop_text = "#DEAAA6"

green_hi_prop = "#38B515"
green_hi_prop_txt = "#186B00"
green_lo_prop_text = "#97BA8B"

cyan_hi_prop = "#00A891"
cyan_hi_prop_txt = "#005E50"
cyan_lo_prop_text = "#85B2AB"

# theme

my_theme <- theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.position = "none",
    axis.text.x = element_text(
      size = 8, hjust = 0.5, color = pink_dark, family = "roboto_slab"
    ),
    axis.text.y = element_text(
      size = 8, hjust = 0.5, color = pink_dark, family = "roboto_slab"
    ),
    axis.title.x = element_text(
      face = "bold", size = 10, vjust = -3, color = pink, family = "Arial"
    ),
    axis.title.y = element_text(
      face = "bold", size = 10, vjust = 3, color = pink, family = "Arial"
    ),
    plot.margin = margin(
      b = 20, l = 20
    ),
    plot.title = element_text(
      size = 15, color = pink, face = "bold", margin = margin(t = 10), 
      family = "Arial Black"
    )
  )

```


# Intersecting Favorability Distributions

```{r}
favorability_levels = c(
    "Highly\nFavorable",
    "Moderately\nFavorable",
    "Ambivalent\nor Neutral",
    "Unfavorable"
  )
```

## Wrangle

Gender Identity

```{r}
heat_trans <- survey %>%
  drop_na(Transphobia_cat, Furry_unfavorability_cat) %>%
  mutate(across(
    c(Transphobia_cat, Furry_unfavorability_cat), 
    \(x) factor(case_match(
      x,
      1 ~ favorability_levels[1],
      2 ~ favorability_levels[2],
      3 ~ favorability_levels[3],
      4 ~ favorability_levels[4]
    ),
    levels = favorability_levels,
    ordered = TRUE)
  )) %>% 
  count(Transphobia_cat, Furry_unfavorability_cat, 
        name = "Count",
        .drop = FALSE) %>% 
  group_by(Transphobia_cat) %>% 
  mutate(Proportion = Count / sum(Count)) %>%
  rename(
    `Attitude on Gender Identity` = Transphobia_cat,
    `Attitude on Furries` = Furry_unfavorability_cat
  )

heat_trans
```

Sexual Orientation

```{r}
heat_gay <- survey %>%
  drop_na(Homophobia_cat, Furry_unfavorability_cat) %>%
  mutate(across(
    c(Homophobia_cat, Furry_unfavorability_cat), 
    \(x) factor(case_match(
      x,
      1 ~ favorability_levels[1],
      2 ~ favorability_levels[2],
      3 ~ favorability_levels[3],
      4 ~ favorability_levels[4]
    ),
    levels = favorability_levels,
    ordered = TRUE)
  )) %>% 
  count(Homophobia_cat, Furry_unfavorability_cat, 
        name = "Count",
        .drop = FALSE) %>% 
  group_by(Homophobia_cat) %>% 
  mutate(Proportion = Count / sum(Count)) %>%
  rename(
    `Attitude on Sexual Orientation` = Homophobia_cat,
    `Attitude on Furries` = Furry_unfavorability_cat
  )

heat_gay
```

## Scrapped: Faceted Histograms

Demonstrated over gender identity

```{r}
enclosure_rectangles <- tibble(
  bottom = unique(survey$Furry_unfavorability_cat) - 0.25,
  top = unique(survey$Furry_unfavorability_cat) + 0.65
) %>%
  drop_na()
```

```{r}
heat_trans %>%
  ggplot(aes(
    x = `Attitude on Furries`,
    y = `Attitude on Gender Identity`
  )) +
  
  # enclosure
  geom_rect(
    data = enclosure_rectangles, 
    aes(
      xmin = -Inf,
      xmax = Inf,
      ymin = bottom,
      ymax = top
    ), 
    fill = pink_faint,
    inherit.aes = FALSE
    ) +
  
  geom_rect(
    aes(
      xmin = as.numeric(`Attitude on Furries`),
      xmax = as.numeric(`Attitude on Furries`) + 1,
      ymin = as.numeric(`Attitude on Gender Identity`),
      ymax = as.numeric(`Attitude on Gender Identity`) + Proportion
    ), fill = pink) +
  
  geom_text(
    aes(
      label = paste0(round(Proportion * 100, digits = 0), "%")),
    nudge_y = -0.2, 
    hjust = 0, vjust = 0, 
    color = pink_mild,
    fontface = "bold", size = 2
  ) +
  
  labs(
    title = "Title"
  ) +
  
  coord_cartesian(xlim = c(1.5, 4.5)) +
  
  my_theme +
  theme(
    axis.text.y = element_text(vjust = 0, hjust = 0),
    axis.text.x = element_text(hjust = 0),
    plot.margin = margin(l = 20, b = 20, r = 30)
  )
```

```{r}
ggsave("scrapped_hist.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

## vs. Gender Identity

Correlation coefficient

```{r}
spearman.test(survey$Furry_unfavorability_cat, survey$Transphobia_cat)
```

```{r}
heat_trans %>%
  ggplot(aes(
    `Attitude on Gender Identity`,
    `Attitude on Furries`,
    fill = Proportion
  )) +
  geom_tile(aes(width = 0.8)) +
  geom_text(
    aes(
      color = Proportion,
      label = paste0(round(Proportion * 100, digits = 0), "%")),
    nudge_x = -0.35, hjust = 0, fontface = "bold", size = 3
  ) +
  
  labs(
    title = "Trans Favorability vs Furry Favorability"
  ) +
  
  scale_fill_gradient(low = lo_prop, high = cyan_hi_prop) +
  scale_color_gradient(low = cyan_lo_prop_text, high = cyan_hi_prop_txt) +
  
  my_theme +
  theme(
    plot.title = element_text(color = cyan, size = 10),
    axis.title.x = element_text(color = cyan)
  )
```

```{r}
ggsave("vs_gender.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

## vs. Sexual Orientation

Correlation coefficient

```{r}
spearman.test(survey$Furry_unfavorability_cat, survey$Homophobia_cat)
```

```{r}
heat_gay %>%
  ggplot(aes(
    `Attitude on Sexual Orientation`,
    `Attitude on Furries`,
    fill = Proportion
  )) +
  geom_tile(aes(width = 0.8)) +
  geom_text(
    aes(
      color = Proportion,
      label = paste0(round(Proportion * 100, digits = 0), "%")),
    nudge_x = -0.35, hjust = 0, fontface = "bold", size = 3
  ) +
  
  labs(
    title = "Gay Favorability vs Furry Favorability"
  ) +
  
  scale_fill_gradient(low = lo_prop, high = green_hi_prop) +
  scale_color_gradient(low = green_lo_prop_text, high = green_hi_prop_txt) +
  
  my_theme +
  theme(
    plot.title = element_text(color = green, size = 10),
    axis.title.x = element_text(color = green)
  )
```

```{r}
ggsave("vs_orient.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

# Prediction Heatmap

## Wrangle

```{r}
predictions <- read_csv("predictions.csv") %>%
  mutate(across(
    c(Actual, Predicted), 
    \(x) factor(case_match(
      x,
      1 ~ favorability_levels[1],
      2 ~ favorability_levels[2],
      3 ~ favorability_levels[3],
      4 ~ favorability_levels[4]
    ),
    levels = favorability_levels,
    ordered = TRUE)
  )) %>% 
  count(Actual, Predicted, 
        name = "Count",
        .drop = FALSE) %>% 
  complete(Actual, Predicted, fill = list(Count = 0)) %>%
  group_by(Actual) %>%
  mutate(Proportion = Count / sum(Count))

predictions
```

```{r}
correct_only <- predictions %>%
  filter(Actual == Predicted)
```

## Heatmap

```{r}
predictions %>%
  ggplot(aes(
    x = Actual,
    y = Predicted,
    fill = Proportion
  )) +
  
  # yellow for correct
  geom_tile(
    data = correct_only,
    aes(
      x = Actual,
      y = Predicted
    ),
    fill = yellow,
    inherit.aes = FALSE
  ) +
  
  geom_tile(aes(width = 0.8)) +
  geom_text(
    aes(
      color = Proportion,
      label = paste0(round(Proportion * 100, digits = 0), "%")),
    nudge_x = -0.35, hjust = 0, fontface = "bold", size = 3
  ) +
  
  labs(
    title = "Furry Favorability Predictions"
  ) +
  
  scale_fill_gradient(low = lo_prop, high = pink_hi_prop) +
  scale_color_gradient(low = pink_lo_prop_text, high = pink_hi_prop_txt) +
  
  my_theme +
  theme(
    plot.title = element_text(size = 14)
  )
```

```{r}
ggsave("true_vs_predicted.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

# Favorability Score Distributions

## Wrangle

```{r}
survey2 <- survey %>%
  drop_na(Homophobia_cat, Furry_unfavorability_cat) %>%
  mutate(across(
    c(
      Furry_unfavorability_cat,
      Homophobia_cat,
      Transphobia_cat
    ), 
    \(x) factor(case_match(
      x,
      1 ~ favorability_levels[1],
      2 ~ favorability_levels[2],
      3 ~ favorability_levels[3],
      4 ~ favorability_levels[4]
    ),
    levels = favorability_levels,
    ordered = TRUE)
  ))
```

## Bar Plots

```{r}
survey2 %>%
  ggplot(aes(Furry_unfavorability_cat)) +
  geom_bar(fill = pink) +
  
  labs(
    title = "Attitude on furries"
  ) +
  
  my_theme + 
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(
      size = 17
    )
  )
```

```{r}
ggsave("dist_furry.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

```{r}
survey2 %>%
  ggplot(aes(Homophobia_cat)) +
  geom_bar(fill = green) +
  
  labs(
    title = "Attitude on sexual orientation"
  ) +
  
  my_theme + 
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(
      size = 17, color = green
    )
  )
```

```{r}
ggsave("dist_orient.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

```{r}
survey2 %>%
  ggplot(aes(Transphobia_cat)) +
  geom_bar(fill = cyan) +
  
  labs(
    title = "Attitude on gender identity"
  ) +
  
  my_theme + 
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(
      size = 17, color = cyan
    )
  )
```

```{r}
ggsave("dist_gender.png", 
       width = 4, height = 3, dpi = my_dpi, bg = "transparent")
```

# All Question Distributions

## Favorability Questions

```{r}
prefixes = c(
  "Furry-", "Homop-", "Trans-"
)
# prefixes

question_names <- colnames(survey)
question_names <- str_subset(
  question_names, paste0("^", prefixes, collapse = "|")
)
# question_names

for (col in question_names) {
  survey %>%
    ggplot(aes(.data[[col]])) +
    geom_bar(fill = pink) +
    
    labs(
      title = col
    ) +
    
    my_theme + 
    theme(
      axis.text.x = element_text(size = 2),
      axis.text.y = element_text(size = 2),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 2),
      plot.title = element_text(
        size = 3, color = pink
      )
    )
  
  ggsave(paste0("question_distributions/", col, ".png"), 
         width = 4, height = 3, dpi = 100, bg = "transparent")
}
```

## Supplemental questions

```{r}
question_names <- c(
  "Perc_guess",
  "Furry_id", "Heterosexual_id", "Transgender+_id",
  "Homophobic_id", "Transphobic_id",
  "Politics",
  "Location", "Medium"
)
# question_names

for (col in question_names) {
  survey %>%
    ggplot(aes(.data[[col]])) +
    geom_bar(fill = pink) +
    
    labs(
      title = col
    ) +
    
    my_theme + 
    theme(
      axis.text.x = element_text(size = 2, angle = 45, hjust = 1),
      axis.text.y = element_text(size = 2),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 2),
      plot.title = element_text(
        size = 3, color = pink
      )
    )
  
  ggsave(paste0("question_distributions/", col, ".png"), 
         width = 4, height = 3, dpi = 100, bg = "transparent")
}
```

# Raw Favorability Distributions

Final favorability score thresholds for visualization of cutoffs
```{r}
# Category 1, Highly favorable : [1 - t1)
t1 = 1.5
# Category 2, Moderately favorable : [t1 - t2)
t2 = 2.5
# Category 3, Ambivalent or neutral : [t2 - t3)
t3 = 3.5
# Category 4, Unfavorable : [t3 - 5]
```

```{r}
survey %>%
  ggplot(aes(Furry_unfavorability_raw)) +
  geom_density(
    bw = 0.15,
    fill = pink, color = pink
  ) +
  
  geom_vline(
    xintercept = c(t1, t2, t3),
    linetype = "dotted",
    color = pink_dark,
    linewidth = 1
  ) +
  
  ylim(0, 1) +
  xlim(1, 5) +
  
  labs(
    title = "Raw Furry"
  ) +
  
  my_theme + 
  theme(
    axis.text.x = element_text(size = 2),
    axis.text.y = element_text(size = 2),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 2),
    plot.title = element_text(
      size = 2.5, color = pink
    )
  )
```

```{r}
ggsave("dist_raw_furry.png", 
       width = 4, height = 3, dpi = 100, bg = "transparent")
```

```{r}
survey %>%
  ggplot(aes(Homophobia_raw)) +
  geom_density(
    bw = 0.15,
    fill = green, color = green
  ) +
  
  geom_vline(
    xintercept = c(t1, t2, t3),
    linetype = "dotted",
    color = pink_dark,
    linewidth = 1
  ) +
  
  ylim(0, 1) +
  xlim(1, 5) +
  
  labs(
    title = "Raw Sexual Orientation"
  ) +
  
  my_theme + 
  theme(
    axis.text.x = element_text(size = 2),
    axis.text.y = element_text(size = 2),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 2),
    plot.title = element_text(
      size = 2.5, color = pink
    )
  )
```

```{r}
ggsave("dist_raw_orient.png", 
       width = 4, height = 3, dpi = 100, bg = "transparent")
```

```{r}
survey %>%
  ggplot(aes(Transphobia_raw)) +
  geom_density(
    bw = 0.15,
    fill = cyan, color = cyan
  ) +
  
  geom_vline(
    xintercept = c(t1, t2, t3),
    linetype = "dotted",
    color = pink_dark,
    linewidth = 1
  ) +
  
  ylim(0, 1) +
  xlim(1, 5) +
  
  labs(
    title = "Raw Gender Identity"
  ) +
  
  my_theme + 
  theme(
    axis.text.x = element_text(size = 2),
    axis.text.y = element_text(size = 2),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 2),
    plot.title = element_text(
      size = 2.5, color = pink
    )
  )
```

```{r}
ggsave("dist_raw_gender.png", 
       width = 4, height = 3, dpi = 100, bg = "transparent")
```

