---
title: "Transform Data"
author: "Mary Krouse"
date: "2026-06-29"
output: html_document
---

# Setup & Imports

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(textclean)
```

```{r}
answers_raw <- read_csv("answers-6-29.csv")
```

## Respondent table

Just a series from 1 to the number of respondents. I can imagine this being useful if any respondents respond to 0 of the questions. Thus far, at least one such respondent exists.

```{r}
table_respondents = tibble(respondent_id = 1:nrow(answers_raw))
table_respondents
```

```{r}
write_csv(table_respondents, 
          "table_respondents.csv")
```

# Tables relating to answers

## Questions table

```{r}
question_id <- c(
  # Furry unfavorability
  "Furry-01", "Furry-02", "Furry-03", "Furry-04", "Furry-05",
  "Furry-06", "Furry-07", "Furry-08", "Furry-09", "Furry-10",
  "Furry-11", "Furry-12", "Furry-13", "Furry-14", "Furry-15",
  "Furry-16", "Furry-17", "Furry-18", "Furry-19", "Furry-20",
  "Furry-21", "Furry-22", "Furry-23", "Furry-24",
  
  # Homophobia
  "Homop-01", "Homop-02", "Homop-03", "Homop-04", "Homop-05",
  "Homop-06", "Homop-07", "Homop-08", "Homop-09", "Homop-10",
  "Homop-11", "Homop-12", "Homop-13", "Homop-14", "Homop-15",
  "Homop-16", "Homop-17", "Homop-18", "Homop-19", "Homop-20",
  
  # Transphobia
  "Trans-01", "Trans-02", "Trans-03", "Trans-04", "Trans-05",
  "Trans-06", "Trans-07", "Trans-08", "Trans-09", "Trans-10",
  "Trans-11", "Trans-12", "Trans-13", "Trans-14", "Trans-15",
  "Trans-16", "Trans-17", "Trans-18", "Trans-19", "Trans-20",
  "Trans-21", "Trans-22",
  
  # Estimated percent of furries LGBTQ+
  "Perc_guess",
  
  # I consider myself or identify as ____
  "Furry_id", "Heterosexual_id", "Transgender+_id", 
  "Homophobic_id", "Transphobic_id", 
  
  # More respondent info
  "Politics", "Location", "Medium"
)

question_text <- colnames(answers_raw)[-(1:2)]

category_number <- c(
  rep(1, times = 24), # 1: Furry unfavorability assessment
  rep(2, times = 20), # 2: Homophobia assessment
  rep(3, times = 22), # 3: Transphobia assessment
  4,                  # 4: Percentage furries LGBTQ+ guess
  rep(5, times = 5),  # 5: Self-identifies as...
  6,                  # 6: Self-described politics
  7,                  # 7: Location where survey was found
  8                   # 8: On what the link to the survey was found
)

table_questions <- tibble(
  question_id = question_id,
  question_text = question_text,
  category_number = category_number
) %>%

  # Curly quotes were causing issues when copying into SQL
  mutate(question_text = str_replace_all(question_text, "[“”‘’]", "'"))

table_questions
```

```{r}
write_csv(table_questions, 
          "table_questions.csv")
```

## Question Category Lookup table

```{r}
category_text = c(
  "Furry unfavorability assessment",
  "Homophobia assessment",
  "Transphobia assessment",
  "Percentage furries LGBTQ+ guess",
  "Self-identifies as...",
  "Self-described politics",
  "Location where survey was found",
  "On what the link to the survey was found"
)

table_category_lookup = tibble(
  category_number = 1:length(category_text),
  category_text = category_text
)

table_category_lookup
```

```{r}
write_csv(table_category_lookup, 
          "table_category_lookup.csv")
```

## Answer Lookup table

```{r}
table_answer_lookup = tribble(
  ~category_number, ~answer_number, ~answer_text,
  
  # 1: Furry unfavorability assessment
  1, 1, "Strongly agree",
  1, 2, "Agree",
  1, 3, "Neither agree nor disagree",
  1, 4, "Disagree",
  1, 5, "Strongly disagree",
  
  # 2: Homophobia assessment
  2, 1, "Strongly agree",
  2, 2, "Agree",
  2, 3, "Neither agree nor disagree",
  2, 4, "Disagree",
  2, 5, "Strongly disagree",
  
  # 3: Transphobia assessment
  3, 1, "Strongly agree",
  3, 2, "Agree",
  3, 3, "Neither agree nor disagree",
  3, 4, "Disagree",
  3, 5, "Strongly disagree",
  
  # 4: Percentage furries LGBTQ+ guess
  4, 1, "0%-10%",
  4, 2, "10%-20%",
  4, 3, "20%-30%",
  4, 4, "30%-40%",
  4, 5, "40%-50%",
  4, 6, "50%-60%",
  4, 7, "60%-70%",
  4, 8, "70%-80%",
  4, 9, "80%-90%",
  4, 10, "90%-100%",
  
  # 5: Self-identifies as...
  5, 1, "Agree",
  5, 2, "Disagree",
  5, 3, "Unsure",
  
  # 6: Self-described politics
  6, 1, "Far left",
  6, 2, "Moderate left",
  6, 3, "Center",
  6, 4, "Moderate right",
  6, 5, "Far right",
  
  # 7: Location where survey was found
  7, 1, "At the Salem Saturday Market (Salem, OR)",
  7, 2, "At the Vashon Island Strawberry Festival (Vashon, WA)", 
  7, 3, "Downtown Salem or Minto Brown Park", 
  7, 4, "Online or sent to me, and I'm taking this survey from within the United States", 
  7, 5, "Online or sent to me, and I'm taking this survey from outside the United States", 
  7, 6, "None of the above", 
  
  # 8: On what the link to the survey was found
  8, 1, "On a tee shirt worn by a student", 
  8, 2, "On a tee shirt worn by a student in a furry costume", 
  8, 3, "On a poster", 
  8, 4, "On a Reddit post", 
  8, 5, "On a LinkedIn post", 
  8, 6, "Sent to me by someone", 
  8, 7, "On SurveySwap.io", 
  8, 8, "None of the above" 
)

table_answer_lookup
```

```{r}
write_csv(table_answer_lookup, 
          "table_answer_lookup.csv")
```

## Respondent-Question-Answer table

```{r}
table_respondent_question_answer <- answers_raw %>%
  
  # Remove unneeded columns
  select(-(1:2)) %>%
  
  # Pivot to long (with index as identifier column)
  rowid_to_column("respondent_id") %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(
    cols = -respondent_id,
    names_to = "question_text",
    values_to = "answer_number"
  ) %>%
  
  # Replace question text with question ID
  left_join(table_questions, by = "question_text") %>%
  select(-question_text) %>%
  
  # Correct text answers that should be answer numbers (beyond category 3)
  left_join(
    table_answer_lookup,
    by = c(
      "category_number" = "category_number",
      "answer_number" = "answer_text"
    )) %>%
  mutate(
    final_answer_id = case_when(
      category_number %in% 1:3 ~ as.integer(answer_number),
      TRUE ~ as.integer(answer_number.y)
    )) %>%
  
  # Delete excess columns
  select(
    -answer_number, 
    -answer_number.y,
    -category_number
  ) %>%
  rename(answer_number = final_answer_id)

table_respondent_question_answer
```

```{r}
write_csv(table_respondent_question_answer, 
          "table_respondent_question_answer.csv")
```

# Table for cleaned raw data with favorability scores

Function for calculating favorability score

```{r}
compute_scale_score <- function(
    df, prefix, exclude_indices = NULL, reverse_indices = NULL) {
  
  # Select specified question category (by prefix)
  cols <- select(df, starts_with(prefix))
  
  # Build a matrix of values for row-wise operations
  vals <- as.matrix(as.data.frame(cols))
  
  # Remove excluded columns first
  if (!is.null(exclude_indices)) {
    vals <- vals[, -exclude_indices, drop = FALSE]
    
    # Shift over indices to replace removed columns
    adjusted_reverse <- sapply(reverse_indices, function(idx) {
      idx - sum(exclude_indices < idx)
    })
    reverse_indices <- adjusted_reverse
  }
  
  # Reverse score specified items: -x + 6
  if (!is.null(reverse_indices)) {
    vals[, reverse_indices] <- -vals[, reverse_indices] + 6
  }
  
  # Compute row means, ignoring NAs
  means = rowMeans(vals, na.rm = TRUE)
  
  # Reverse once again to match up with favorability score categories
  -means + 6
}
```

Thresholds for favorability category assignment by raw score

```{r}
# Category 1, Highly favorable : [1 - t1)
t1 = 1.5
# Category 2, Moderately favorable : [t1 - t2)
t2 = 2.5
# Category 3, Ambivalent or neutral : [t2 - t3)
t3 = 3.5
# Category 4, Unfavorable : [t3 - 5]
```

```{r}
answers <- answers_raw %>%
  # Remove unneeded columns
  # - 1: Timestamp of response submission (anonymity concerns)
  # - 2: Agreement to consent form (same value for all)
  select(-(1:2)) %>%
  
  # Rename columns to question names
  rename_with(~question_id) %>%
  
  # Get raw favorability scores
  mutate(
    Furry_unfavorability_raw = compute_scale_score(
      ., 
      prefix = "Furry-",
      exclude_indices = c(1, 2, 3, 4),
      reverse_indices = c(5, 10, 16, 19, 21)
    ) * -1 + 6,
    Homophobia_raw = compute_scale_score(
      .,
      prefix = "Homop-",
      exclude_indices = NULL,
      reverse_indices = c(5, 10, 15, 16)
    ) * -1 + 6,
    Transphobia_raw = compute_scale_score(
      .,
      prefix = "Trans-",
      exclude_indices = NULL,
      reverse_indices = c(1, 2, 5, 6, 8, 9, 10, 11, 21)
    )) %>%
  
  mutate(
    Furry_unfavorability_cat = as_factor(case_when(
      Furry_unfavorability_raw < t1 ~ 1,
      Furry_unfavorability_raw < t2 ~ 2,
      Furry_unfavorability_raw < t3 ~ 3,
      Furry_unfavorability_raw >= t3 ~ 4
    )),
    
    Homophobia_cat = as_factor(case_when(
      Homophobia_raw < t1 ~ 1,
      Homophobia_raw < t2 ~ 2,
      Homophobia_raw < t3 ~ 3,
      Homophobia_raw >= t3 ~ 4
    )),
    
    Transphobia_cat = as_factor(case_when(
      Transphobia_raw < t1 ~ 1,
      Transphobia_raw < t2 ~ 2,
      Transphobia_raw < t3 ~ 3,
      Transphobia_raw >= t3 ~ 4
    )))

answers
```

```{r}
write_csv(answers, 
          "answers_clean.csv")
```

# Tables relating to favorability scores

## Favorability Lookup table

```{r}
fav_description = c(
  "Highly favorable",
  "Moderately favorable",
  "Ambivalent or neutral",
  "Unfavorable"
)

table_favorability_lookup <- tibble(
  fav_number = 1:length(fav_description),
  fav_description = fav_description
)

table_favorability_lookup
```

```{r}
write_csv(table_favorability_lookup, 
          "table_favorability_lookup.csv")
```

## Respondent Favorability table

```{r}
table_respondent_favorability <- answers %>%
  # Isolate favorability score columns
  rowid_to_column("respondent_id") %>%
  select(
    respondent_id,
    Furry_unfavorability_cat,
    Homophobia_cat,
    Transphobia_cat
  ) %>%
  
  # Pivot long with names of groups and favorability toward each
  rename(
    Furry = Furry_unfavorability_cat,
    Gay = Homophobia_cat,
    Trans = Transphobia_cat
  ) %>%
  pivot_longer(
    cols = -respondent_id,
    names_to = "towards_group",
    values_to = "fav_number"
  )

table_respondent_favorability
```

```{r}
write_csv(table_respondent_favorability, 
          "table_respondent_favorability.csv")
```
